### 2020-02-27 dfol_draft/defol_draft2.Rmd

---
title: "*dfoliatR*: An R package for identification and analyses of insect defoliation and growth suppression signals in tree rings"
authors:
  - Christopher H. Guiterman:
      email: chguiterman@email.arizona.edu
      affiliation: [LTRR]
      correspondence: true
    # footnote: 1
  - Ann M. Lynch:
      email: lyncha@email.arizona.edu
      affiliation: [RMRS, LTRR]
      correspondence: false
  - Jodi Axelson:
      email: jodi.axelson@berkeley.edu 
      affiliation: [UCB]
      correspondence: false
institute:
  - LTRR: Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - RMRS: U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - UCB: Department of Environmental Science, Policy & Management, University of California, Berkeley, 120 Mulford Hall, Berkeley, CA
abstract: For author review only. **Please do not share**
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  bookdown::word_document2:
    reference_docx: "../common/word-styles-reference-01.docx"
    toc: false
    toc_depth: 2
    fig_caption: yes
    df_print: kable
bibliography: "../common/Report.bib"
csl: "../common/ecology.csl"
keep_md: false
link-citations: false
---
```{r setup, comment=FALSE, message = FALSE, echo=FALSE, warning=FALSE}
rm(list=ls())           # Clean the environment
options(scipen=6)       # display digits proberly!! not the scientific version
options(digits.secs=6)  # use milliseconds in Date/Time data types
options(warning=FALSE)  # don't show warnings
options(figure_counter = TRUE)
library(knitr)          # set global knitr options of the document
library(dplyr)
# Here we set the R code invisible to not show when the document is rendered
opts_chunk$set(comment="", message = FALSE, echo=FALSE, error=FALSE, warning=FALSE,
               fig.show='asis', dev='pdf')
```

# Introduction

Some text.