### 2020-02-27 dfol_draft/dfol_draft.Rmd

---
title: "*dfoliatR*: An R package for identification and analyses of insect defoliation and growth suppression signals in tree rings"
author:
  - Christopher H. Guiterman:
      email: chguiterman@email.arizona.edu
      affiliation: [LTRR]
      correspondence: true
    # footnote: 1
  - Ann M. Lynch:
      email: lyncha@email.arizona.edu
      affiliation: [RMRS, LTRR]
      correspondence: false
  - Jodi Axelson:
      email: jodi.axelson@berkeley.edu 
      affiliation: [UCB]
      correspondence: false
    # footnote: 2
# address:
#   - code: LTRR
#     address: Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
#   - code: RMRS
#     address: U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
#   - code: UCB
#     address: Department of Environmental Science, Policy & Management, 120 Mulford Hall, University of California, Berkeley
# Berkeley, CA 94720
institute:
  - LTRR: Laboratory of Tree-Ring Research, University of Arizona, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - RMRS: U.S. Forest Service, Rocky Mountain Research Station, 1215 E Lowell St. Box 210045, Tucson, AZ, 85721
  - UCB: Department of Environmental Science, Policy & Management, University of California, Berkeley, 120 Mulford Hall, Berkeley, CA 94720
# footnote:
#   - code: 1
#     text: "Corresponding Author (chguiterman@email.arizona.edu)"
#   # - code: 2
#   #   text: "Equal contribution"
abstract: |
  This is the abstract.

  It consists of two paragraphs.
keywords: "Defoliation, Insect outbreak"
journal: "Tree-Ring Research"
date: "`r Sys.Date()`"
# bibliography: mybibfile.bib
# linenumbers: true
#numbersections: true
# csl: elsevier-harvard.csl
# output: rticles::elsevier_article
# output: 
#     word_document:
#       toc: no
#       pandoc_args:
#         - '--lua-filter=scholarly-metadata.lua'
#         - '--lua-filter=author-info-blocks.lua'
#     pdf_document:
#       rticles::elsevier_article
#       pandoc_args:
#         - '--lua-filter=scholarly-metadata.lua'
#         - '--lua-filter=author-info-blocks.lua'
# output: rticles::elsevier_article
output:
 bookdown::word_document2:
   # reference_docx: "../extras/word-styles-reference-01.docx"
   # toc: false
   # toc_depth: 2
   # fig_caption: yes
   # df_print: kable
    pandoc_args:
      - --csl=Extras/clinical-mass-spectrometry.csl
      - --citation-abbreviations=Extras/abbreviations.json
      # - --filter=pandoc-crossref
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
      - --reference-doc=Extras/Reference_Document.docx 
bibliography: "Extras/mybibfile.bib"
toc: false
---

_Text based on elsarticle sample manuscript, see [http://www.elsevier.com/author-schemas/latex-instructions#elsarticle](http://www.elsevier.com/author-schemas/latex-instructions#elsarticle)_

The Elsevier article class
==========================

#### Installation

If the document class *elsarticle* is not available on your computer,
you can download and install the system package *texlive-publishers*
(Linux) or install the LaTeX package *elsarticle* using the package
manager of your TeX installation, which is typically TeX Live or MikTeX.

#### Usage

Once the package is properly installed, you can use the document class
*elsarticle* to create a manuscript. Please make sure that your
manuscript follows the guidelines in the Guide for Authors of the
relevant journal. It is not necessary to typeset your manuscript in
exactly the same way as an article, unless you are submitting to a
camera-ready copy (CRC) journal.

#### Functionality

The Elsevier article class is based on the standard article class and
supports almost all of the functionality of that class. In addition, it
features commands and options to format the

-   document style

-   baselineskip

-   front matter

-   keywords and MSC codes

-   theorems, definitions and proofs

-   lables of enumerations

-   citation style and labeling.

Front matter
============

The author names and affiliations could be formatted in two ways:

(1) Group the authors per affiliation.

(2) Use footnotes to indicate the affiliations.

See the front matter of this document for examples. You are recommended
to conform your choice to the journal you are submitting to.

Bibliography styles
===================

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use BibTeX to generate your bibliography and include DOIs
whenever available.

Here are two sample references: @Feynman1963118 [@Dirac1953888].

References {#references .unnumbered}
==========