### 2020-02-27 paper/paper.Rmd

---
title: Title here
subtitle: Do you have a subtitle? If so, write it here
titlerunning: Short form of title (if too long for head)
authorrunning: Short form of author list if too long for running head
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

authors: 
- name: Äüthör 1
  address: Department of YYY, University of XXX
  email: abc@def

- name: Âuthóř 2
  address: Department of ZZZ, University of WWW
  email: djf@wef

keywords:
- key
- dictionary
- word

#PACS: 
#- PAC1
#- superPAC

MSC:
- MSC code 1
- MSC code 2

abstract: |
  The text of your abstract.  150 -- 250 words.

bibliography: bibliography.bib
bibstyle: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
output: rticles::springer_article
---

# Introduction {#intro}

Your text comes here. Separate text sections with \cite{Mislevy06Cog}.

# Section title {#sec:1}

Text with citations by \cite{Galyardt14mmm}.

## Subsection title {#sec:2}

as required. Don't forget to give each section
and subsection a unique label (see Sect. \ref{sec:1}).

#### Paragraph headings

Use paragraph headings as needed.

\begin{align}
a^2+b^2=c^2
\end{align}