### 2020-05-21 README.Rmd
update readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings

<!-- badges: start -->
<!-- badges: end -->

A repository for the initial descriptive paper of the [`dfoliatR`](https://chguiterman.github.io/dfoliatR/) package in R. The paper is currently being prepared for submission to *Dendrochronologia*.

Please contact Chris Guiterman with questions or comments.