### 2020-09-02 README.Rmd
updates for proof and publication

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### `dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings

<!-- badges: start -->
<!-- badges: end -->

This repository includes the code and text to generate a preprint version of the published paper:

>Guiterman, CH, AM Lynch, and JN Axelson (2020) [`dfoliatR`: An R package for detection and analysis of insect defoliation signals in tree rings.](https://authors.elsevier.com/sd/article/S1125786520300898) _Dendrochronologia_. DOI: 10.1016/j.dendro.2020.125750.

The paper provides an overview of the `dfoliatR` R package. More information on using the software is availabe from [https://chguiterman.github.io/dfoliatR/](https://chguiterman.github.io/dfoliatR/).

Please contact @chguiterman with questions or comments.